Comprehensive Quality-Aware Automated Semantic Web Service Composition

### 2017-03-16 paper.tex
update

\begin{document}
\maketitle
\begin{abstract}
Semantic web service composition has been a prevailing research area in recent years. There are two major challenges faced by researchers, semantic matchmaking and QoS optimisation. Semantic matchmaking aims to discover interoperable web services that can interact with each other to provide rich functionalities through language understanding and reasoning. QoS optimisation aims to optimise the non-functional requirement of service users, e.g., minimum cost, maximal reliability. Many scholars have looked into QoS optimisation problems in QoS-aware web service composition applying AI planning and Evolutionary Computation techniques. To meet users' requirement, one often needs to consider both semantic matchmaking quality and QoS simultaneously. Existing works on web service composition often concerns either semantic web  service composition or QoS-aware web service composition. However, there are few works concerns both semantic matchmaking quality and QoS at the same time and achieve a more desirable tradeoff in consideration of both sides. Therefore, In this paper, proposed a general comprehensive quality model considering semantic matchmaking quality and QoS. This model is also suitable to be tackled through EC algorithms. And we develop a PSO-based service composition approach with explicit support for the comprehensive model, this method can generate a more desirable balance in considering semantic matchmaking quality and QoS requirement that is evidently supported by our experiment for comparing QoS-aware approach. We also conduct another experiment shows that our PSO-based method performs better in finding more optimised solution by comparing one existing GP-based method.

@@

\textit{Web service composition} pertains to a combination of multiple web services to provide a value-added composition service that accommodates customers' arbitrarily complex requirements. This application is developed by integrating interoperable and collaborative functionalities over heterogeneous systems. Due to the increasing number of large-scale enterprise applications, the number of Web services has increased substantially and unprecedentedly. Therefore, manual and semi-automated web service composition are considered to be less efficient while automated web service composition enjoys less human intervention, less time consumption, and high productivity.

Two most notable challenges for web service composition are $(1)$ ensuring interoperability of services and $(2)$ achieving QoS optimisation \cite{fensel2011semantic}. \textit{Interoperability} of web services presents challenge $(1)$ in syntactic and semantics dimensions. The syntactic dimension is covered by the XML-based technologies (such as $WSDL$, $SOAP$). The semantics aspect, on the other hand, demands further research. Through ontology-based semantics \cite{o2005review}, web services can understand and better collaborate with each other. There are many ontology languages and formats for semantic service descriptions, such as OWL-S, WSML, and SAWSDL \cite{petrie2016web}, which make ``machine understanding'' possible through identifying and matching semantic similarity in input/output parameters of web services in heterogeneous environments. The second challenge $(2)$ is related to finding \textit{optimised solutions} to Quality of Service ($QoS$). This problem gives birth to \textit{QoS-aware service composition} that considers the composition of service-level agreements (SLA) \cite {sahai2002automated} involving a collection of SLA rules and policies for supporting QoS-based composition.

Existing works on service composition focus mainly on addressing the one of challenges above. A lot of works have been conducted to optimise the quality of compositions under a pre-defined abstract workflow, which is generally considered as \textit{semi-automated Web service composition} approach \cite{parejo2008qos,bahadori2009optimal}. Meanwhile, many research works consider the possibility of generating a composition plan automatically in discovering and selecting suitable web services, which are considered to be NP-hard \cite{moghaddam2014service}. \textit{Semantic web services composition} is distinguished from the syntactic service composition, with the hope of eliminating conflicts by the semantic level of web services. In the past few years, substantial works have been done on semantic web service composition \cite{boustil2014semantic, mier2015integrated, bansal2016generalized}. However, few works have enabled truely automatic semantic web service composition, where both QoS and quality of semantic match making will be optimised simultaneously and achieved a desired balance.

The overall goal of this paper is to \textit{develop an general comprehensive approach to automated QoS-aware semantic web service composition that satisfactorily optimises both QoS and semantic matchmaking quality}. Particularly, this paper extends existing works of QoS-aware service composition by considering both QoS optimisation and semantic matchmaking quality optimisation in our proposed comprehensive quality model. Particle Swarm Optimisation (PSO) has show its promise in searching near-optimised service composition solutions \cite{da2016particle}. We will propose a PSO-based service composition approach with explicit support for our proposed quality mode. We will achieve three objectives in this work as follows:

\begin{enumerate}
 \item To propose a general comprehensive quality model that address QoS and semantic matchmaking quality in considering different matching types with corresponding concept similarity.

@@

 \item To evaluate our proposed approach, we conduct experiment to compare our comprehensive quality model with one widely used QoS evaluation model, we also compare an one existing GP approach with our PSO-based approach using our proposed model. Both comparison utilise benchmark datasets from Web Services Challenge 2009 (WSC09) \cite{kona2009wsc}

@@

The aim of web service composition is to fulfil customers' requirements, which could be denoted as a composition goal: $Comp.G(F(I_{T}, O_{T}), NF(QoS_{T}))$. This overall composition goal is demonstrated in two parts: functional part and non-functional part. The functional part is defined by given task input and desired task output. The nonfunctional part specifies overall QoS requirements. To accomplish the composition goal, two stages are involved, services discovery and service selection. Service discovery is to find matching web services: $S_{n}(F(I_{S}, O_{S}), NF(QoS_{S}))$ from a Service Repository: $\mathcal{S}$ $=  \{S_{1}, S_{2},..., S_{n} \}$ for the given $I_{T}$. If no atomic web service could satisfy the composition goal, a composition of web services will be found to meet $Comp.G$. To ensure the composition solution returns the desired $Comp.G$, we use a quality model that take care of QoS and semantic matchmaking quality.

\subsection{Quality of Service and Composition Constructs}\label{Quality of Service and Composition Constructs}
According to \cite{zeng2003quality}, four most often considered QoS parameters are response time, cost, reliability and availability described as follows:
\begin{itemize}
\item \textit{Response time} ($T$) measures the expected delay in seconds between the moment when a request is sent and the moment when the results are received.
\item \textit{Cost} ($C$) is the amount of money that a service requester has to pay for executing the web service
\item \textit{Reliability} ($R$) is the probability that a request is correctly responded within the maximum expected time frame.
\item \textit{Availability} ($A$) is the probability that a web service is accessible.
\end{itemize}
The aggregation value of QoS attributes for a composite service varies with respect to different constructs, which reflects how services associated with each other in a service composition \cite{zeng2003quality}. Here we consider two composition constructs: sequence and parallel constructs, in building the composition web service. The QoS calculation models are described as follows:
\subsubsection{Sequence construct}
The composition web service executes each atomic service associated with a sequence construct in a definite sequence order. The aggregation value for total time ($T$) and total cost ($C$) is as the sum of time and cost of web services involved respectively. The overall availability and reliability in a sequence construct are calculated by multiplying their corresponding availability and reliability of each web service in probability theory. This construct is shown in Fig. \ref{sequence}.
\subsubsection{Parallel construct}
Web services in a parallel construct are executed concurrently. The QoS aggregation value for total cost, availability and reliability are the same as these in sequence construct while the Total time ($T$) is determined by the most time-consuming path in the composition flow of the solution. This construct is presented in Fig. \ref{parallel}.

\section{Comprehensive Quality Model}\label{qswsc_approach}
In this section, we propose a general comprehensive quality-aware service model in consideration of both semantic matchmaking quality and QoS for semantic web service composition.

@@

@@

\begin{table}[]
\centering
\caption{Mean fitness results for comparing GP-based approach}
\label{meanFitness}
\begin{tabular}{l|l|l}
\hline
\multicolumn{1}{c|}{Dataset} & PSO-based approach & GP-based approach  \\ \hline
WSC09 01                     &0.559207 $\pm$ 0.012780 $\uparrow$      &0.518411 $\pm$ 0.018470               \\ \hline
WSC09 02                     &0.470083 $\pm$  0.001106   &0.471594 $\pm$  0.002436 $\uparrow$           \\ \hline
WSC09 03                     & 0.559207 $\pm$ 0.012780 $\pm$ $\uparrow$  &0.552947 $\pm$ 0.007094            \\ \hline
WSC09 04                     & 0.468942 $\pm$ 0.001670 $\pm$ $\uparrow$  &0.442934 $\pm$ 0.010154            \\ \hline
WSC09 05                     & 0.469424 $\pm$ 0.000800 $\pm$ $\uparrow$  &0.256425 $\pm$ 0.038916            \\ \hline
\end{tabular}
\end{table}

>>>>

\begin{document}
\maketitle
\begin{abstract}
Semantic web service composition has been a prevailing research area in recent years. There are two major challenges faced by researchers, semantic matchmaking and Quality of Service (QoS) optimisation. Semantic matchmaking aims to discover interoperable web services that can interact with each other to provide rich functionalities through language understanding and reasoning. QoS optimisation aims to optimise the non-functional requirements of service users, e.g., minimum cost, maximal reliability. Many scholars have looked into QoS optimisation problems in QoS-aware web service composition, applying AI planning and Evolutionary Computation techniques. To meet users' requirements, one often needs to consider both semantic matchmaking quality and QoS simultaneously. Existing works on web service composition often concern either semantic web  service composition or QoS-aware web service composition. However, there are few works concerning both semantic matchmaking quality and QoS at the same time and achieving a more desirable tradeoff in consideration of both sides. Therefore, we propose a general comprehensive quality model considering semantic matchmaking quality and QoS in this paper. This model is also suitable to be tackled through EC algorithms. Further more, we develop a PSO-based service composition approach with explicit support for the comprehensive model. This method can generate a more desirable balance in considering semantic matchmaking quality and QoS requirements that is supported by our experiment when compared to QoS-aware approach. We also conduct another experiment that shows our PSO-based method performs better in finding a more optimised solution by comparing with one existing GP-based method.

@@

\textit{Web service composition} pertains to a combination of multiple web services to provide a value-added composition service that accommodates customers' arbitrarily complex requirements. This application is developed by integrating interoperable and collaborative functionalities over heterogeneous systems. Due to the increasing number of large-scale enterprise applications, the number of Web services has increased substantially and unprecedentedly. Therefore, manual and semi-automated web service compositions are considered to be less efficient while automated web service composition has less human intervention, less time consumption, and high productivity.

Two most notable challenges for web service composition are ensuring interoperability of services and achieving Quality of Service (QoS) optimisation \cite{fensel2011semantic}. \textit{Interoperability} of web services presents challenge in syntactic and semantic dimensions. The syntactic dimension is covered by the XML-based technologies (such as $WSDL$, $SOAP$). The semantic aspect, on the other hand, demands further research. Through ontology-based semantics \cite{o2005review}, web services can understand and better collaborate with each other. There are many ontology languages and formats for semantic service descriptions, such as OWL-S, WSML, and SAWSDL \cite{petrie2016web}, which make ``machine understanding'' possible through identifying and matching semantic similarities in input/output parameters of web services in heterogeneous environments. The second challenge $(2)$ is related to finding \textit{optimised solutions} to QoS. This problem gives birth to \textit{QoS-aware service composition} that considers the composition of service-level agreements (SLA) \cite {sahai2002automated} involving a collection of SLA rules and policies for supporting QoS-based composition.

Existing works on service composition focus mainly on addressing one of challenges above. Many works have been conducted to optimise the quality of compositions under a pre-defined abstract workflow, which is generally considered as a \textit{semi-automated Web service composition} approach \cite{parejo2008qos,bahadori2009optimal}. Meanwhile, many research works consider the possibility of generating a composition plan automatically in discovering and selecting suitable web services, which are considered to be NP-hard \cite{moghaddam2014service}. \textit{Semantic web services composition} is distinguished from the syntactic service composition, with the hope of eliminating conflicts by the semantic level of web services. In the past few years, substantial works have been done on semantic web service composition \cite{boustil2014semantic, mier2015integrated, bansal2016generalized}. However, few works have enabled truely automatic semantic web service composition, where both QoS and quality of semantic match making will be optimised simultaneously and achieved a desired balance.

The overall goal of this paper is to \textit{develop an general comprehensive approach to automated QoS-aware semantic web service composition that satisfactorily optimises both QoS and semantic matchmaking quality}. Particularly, this paper extends existing works of QoS-aware service composition by considering both QoS optimisation and semantic matchmaking quality optimisation in our proposed comprehensive quality model. Particle Swarm Optimisation (PSO) has shown its promise in searching for near-optimised service composition solutions \cite{da2016particle}. We will propose a PSO-based service composition approach with explicit support for our proposed quality mode. We will achieve three objectives in this work as follows:

\begin{enumerate}
 \item To propose a general comprehensive quality model that addresses QoS and semantic matchmaking quality in considering different matching types with corresponding concept similarities.

@@

 \item To evaluate our proposed approach, we conduct experiments to compare our comprehensive quality model with one widely used QoS evaluation model. We also compare one recent GP approach \cite{ma2015hybrid} with our PSO-based approach using our proposed model. Both comparisons utilise benchmark datasets from Web Services Challenge 2009 (WSC09) \cite{kona2009wsc}

@@

The aim of web service composition is to fulfil customers' requirements, which could be denoted as a composition goal: $Comp.G(F(I_{T}, O_{T}), NF(QoS_{T}))$. This overall composition goal is demonstrated in two parts: functional part and non-functional part. The functional part is defined by given task inputs and desired task outputs. The nonfunctional part specifies overall QoS requirements. To accomplish the composition goal, two stages are involved, services discovery and service selection. Service discovery is to find matching web services: $S_{n}(F(I_{S}, O_{S}), NF(QoS_{S}))$ from a Service Repository: $\mathcal{S}$ $=  \{S_{1}, S_{2},..., S_{n} \}$ for the given $I_{T}$. If no atomic web service could satisfy the composition goal, a composition of web services will be found to meet $Comp.G$. To ensure the composition solution returns the desired $Comp.G$ that considers a higher degree of matching $O_{T}$ and a near-optimised $QoS_{T}$, We first demonstrate the concepts in both semantic matchmaking types and $QoS$ as follows, and then propose a quality model that take care of semantic matchmaking quality and QoS simultaneously in Sect. \ref{Comprehensive_Quality_Model}.

@@

\subsection{Quality of Service}\label{Quality of Service and Composition Constructs}
According to \cite{zeng2003quality}, four most often considered QoS parameters are response time, cost, reliability and availability described as follows:
\begin{itemize}
\item \textit{Response time} ($T$) measures the expected delay in seconds between the moment when a request is sent and the moment when the results are received.
\item \textit{Cost} ($C$) is the amount of money that a service requester has to pay for executing the web service
\item \textit{Reliability} ($R$) is the probability that a request is correctly responded within the maximum expected time frame.
\item \textit{Availability} ($A$) is the probability that a web service is accessible.
\end{itemize}
The aggregation value of QoS attributes for a composite service varies with respect to different constructs, which reflects how services associated with each other in a service composition \cite{zeng2003quality}. Here we consider two composition constructs: sequence and parallel constructs, in building the composition web service. The QoS calculation models are described as follows:
\subsubsection{Sequence construct}
all involved atomic services associated with a sequence construct are executed in a definite sequence order. The aggregation value for total time ($T$) and total cost ($C$) is as the sum of time and cost of web services involved respectively. The overall availability and reliability in a sequence construct are calculated by multiplying their corresponding availability and reliability of each web service in probability theory. This construct is shown in Fig. \ref{sequence}.
\subsubsection{Parallel construct}
Web services in a parallel construct are executed concurrently. The QoS aggregation value for total cost, availability and reliability are the same as these in sequence construct while the Total time ($T$) is determined by the most time-consuming path in the composition flow of the solution. This construct is presented in Fig. \ref{parallel}.

\section{Comprehensive Quality Model}\label{Comprehensive_Quality_Model}
In this section, we propose a general comprehensive quality-aware service model in consideration of both semantic matchmaking quality and QoS for semantic web service composition.

@@

\begin{table}[]
\centering
\caption{Mean fitness results for comparing GP-based approach}
\label{meanFitness}
\begin{tabular}{l|l|l}
\hline
\multicolumn{1}{c|}{Dataset} & PSO-based approach & GP-based approach  \\ \hline
WSC09 01                     &0.559207 $\pm$ 0.012780  $\uparrow$ (P:0.000)  &0.520659 $\pm$ 0.020758           \\ \hline
WSC09 02                     &0.470083 $\pm$  0.001106 $\uparrow$ (P:0.008)   &0.459681 $\pm$  0.002874          \\ \hline
WSC09 03                     & 0.550396 $\pm$ 0.012780                        &0.567915 $\pm$ 0.023447 $\uparrow$  (P:0.000) \\ \hline
WSC09 04                     & 0.468942 $\pm$ 0.001670  $\uparrow$ (P:0.003)  &0.431704 $\pm$ 0.009774            \\ \hline
WSC09 05                     & 0.469424 $\pm$ 0.000800  $\uparrow$ (P:0.000)  &0.245222 $\pm$ 0.036885            \\ \hline
\end{tabular}
\end{table}
